Community-Conditioned Language Models

### 2020-11-22 paper/paper.tex
colors for all the community-level plots

%
% File acl2020.tex
%
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

@@

\begin{table*}
  \normalsize
  \centering
  \input{floats/community_results.tex}
  \caption{
    Model performance by community. We show the absolute perplexity 
    of the baseline models and the information gain of the CCLMs.
    Note that the rows are sorted by the perplexity of the baseline
    LSTM model.
}
  \label{tab:community-results}
\end{table*}

@@

\begin{figure*}
\begin{tikzpicture}
  \commpcagroupplot{floats/lstm-3-1_pca.csv}{floats/transformer-3-3_pca.csv}
\end{tikzpicture}
\caption{Community embedding PCA of the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
\label{fig:comm-pca}
\end{figure*}

@@

\begin{figure}
  \pairwisecommsim{lstm-3-1} 
  \caption{Pairwise cosine similarity between community vectors for 
    the membership-based embedding (x-axis) and
  the LSTM CCLM embedding with $c=1$ (y-axis). 
  Pairs from the same category are colored according to \cref{fig:comm-pca}.
  We also mark pairs from different (\texttimes) and pairs where one or both
  communities belong to the ``Other'' category (\textbigcircle).
  }
  \label{fig:pairwise-comm-sim}
\end{figure}

@@

\begin{figure*}
%\begin{tikzpicture}
  %\confusiongroupplot{lstm-3-1}{transformer-3-3}
%\end{tikzpicture}
\caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
\label{fig:confusion}
\end{figure*}

@@

\begin{figure}
  \pplscatter{lstm-3-1_lmcc}{transformer-3-3_lmcc}
  \caption{%
    Correlation between indiscernibility for the best LSTM ($c = 1$) and transformer ($c = 3$) models
    (Pearson's $r = 0.99$, $p < 0.0001$).
  }
  \label{fig:lmcc_ppl}
\end{figure}

@@

\begin{figure}
  \pplscatter{lstm-3-1_cclm}{lstm-3-1_lmcc}
  \caption{%
    Mean community-wise linguistic perplexity (x-axis) 
    and indiscernibility (y-axis) 
    (Pearson's r = \num{0.16}; p = \num{0.2991}).  
    Shown here for the LSTM with $c = 1$.}
  \label{fig:cclm_lmcc_ppl}
\end{figure}

@@

On the other hand, community indiscernibility is correlated with
the information gain of the CCLM 
on messages from that community's test set 
(see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm_lmcc_ppl}).
Communities for which the language model bennifits most from 
the conditioned model are generally easier for the LMCC to classify.

\begin{figure}
  \pplscatter{lstm-3-1_gain}{lstm-3-1_lmcc}
  \caption{%
    Mean community-wise information gain (x-axis) 
    and indiscernibility (y-axis) 
    %(Pearson's r = \num{0.16}; p = \num{0.2991}).  
    Shown here for the LSTM with $c = 1$.}
  \label{fig:cclm_lmcc_ppl}
\end{figure}

>>>>

%

% File acl2020.tex
%
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

@@

\begin{table*}
  \small
  \centering
  \input{floats/community_results.tex}
  \caption{
    Model performance by community. We show the absolute perplexity 
    of the baseline models and the information gain of the CCLMs.
    Note that the rows are sorted by the perplexity of the baseline
    LSTM model.
}
  \label{tab:community-results}
\end{table*}

@@

\begin{figure*}
\begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1}
  ]
    \nextgroupplot
    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
    \nextgroupplot
    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
  \end{groupplot}
\end{tikzpicture}
\caption{Community embedding PCA of the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
\label{fig:comm-pca}
\end{figure*}

@@

\begin{figure}
  \pairwisecommsim{lstm-3-1} 
  \caption{Pairwise cosine similarity between community vectors for 
    the membership-based embedding (x-axis) and
  the LSTM CCLM embedding with $c=1$ (y-axis). 
  Pairs from the same category are colored according to \cref{fig:comm-pca}.
  We also mark pairs from different categories (\texttimes) and pairs where one or both
  communities belong to the ``Other'' category (\textbigcircle).
  }
  \label{fig:pairwise-comm-sim}
\end{figure}

@@

\begin{figure*}
\begin{tikzpicture}
  \confusiongroupplot{lstm-3-1}{transformer-3-3}
\end{tikzpicture}
\caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
\label{fig:confusion}
\end{figure*}

@@

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
    \commscatter{lstm-3-1_indisc}{transformer-3-3_indisc}
    \end{axis}
  \end{tikzpicture}
  \caption{%
    Correlation between indiscernibility for 
    the best LSTM ($c = 1$) and transformer ($c = 3$) models
    (Pearson's $r = 0.99$, $p < 0.0001$).
  }
  \label{fig:lmcc_ppl}
\end{figure}

@@

\begin{figure*}
  \begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    yticklabels={,,}, xticklabels={,,},
    legend style={anchor=north, legend columns=1, font=\tiny},
    every node near coord/.append style={font=\sffamily\tiny}%
  ]
    \nextgroupplot
    \commscatter[]{lstm-3-1_lm_ppl}{lstm-3-1_indisc}
    \nextgroupplot
    \commscatter[]{lstm-3-1_info_gain}{lstm-3-1_indisc}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
  \end{tikzpicture}
  \caption{%
    Mean community indiscernibility (y-axis), 
    in relationship to linguistic complexity (left) and 
    CCLM information gain (right). Shown here for the LSTM
    model with $l_c=1$. See \cref{tab:ind-corrs} for 
    Pearson's r correlations for all models.
  }
  \label{fig:cclm_lmcc_ppl}
\end{figure*}

@@

On the other hand, community indiscernibility is correlated with
the information gain of the CCLM 
on messages from that community's test set 
(see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm_lmcc_ppl}).
Communities for which the language model benefits most from 
the conditioned model are generally easier for the LMCC to classify.

### 2020-11-22 paper/plot_macros.tex
info gain table/plot
partial refactor of tables
colors for all the community-level plots

\newcommand{\commpcaplot}[1]{%
  \addplot[
    scatter, only marks,
    font=\tiny,
    scatter src=explicit symbolic,
    scatter/classes={
      general-interest={mark=*, blue},
      videogames={mark=*, red},
      technology={mark=*, green},
      sports={mark=*, cyan},
      female-focused={mark=*, magenta},
      other={mark=o, black}%
    },
    nodes near coords*={\community},
    visualization depends on={value \thisrow{community} \as \community}
  ] table [x=pca1, y=pca2, meta=category] {#1};
}

\newcommand{\commpcagroupplot}[2]{%
  \begin{groupplot}[
    group style={group size=2 by 1}, 
    yticklabels={,,}, xticklabels={,,},
    legend style={anchor=north, legend columns=1, font=\tiny},
    every node near coord/.append style={font=\sffamily\tiny}%
    ]
    \nextgroupplot
    \commpcaplot{#1}
    \nextgroupplot
    \commpcaplot{#2}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
}

\newcommand{\confusionplot}[1]{%
    \addplot[
        matrix plot*,
        shader=faceted,
        mesh/cols=46,
        point meta=explicit,
] table [x=actual_comm, y=confered_to, meta=avg_prob] {#1};
}

\newcommand{\confusiongroupplot}[2]{%
  \begin{groupplot}[
  group style={group size=2 by 1}, 
  enlargelimits=false,
  axis line style={draw=none},
  %axis equal,
  xmin=-1, xmax=46,
  ymin=-1, ymax=46,
  width=8cm,
  xticklabels={Advice,AskWomen,BabyBumps,CFB,Drugs,EDH,EarthPorn,Fantasy,GameDeals,GlobalOffensive,Jokes,Kappa,KerbalSpaceProgram,KotakuInAction,LifeProTips,MLS,MMA,MaddenUltimateTeam,TwoXChromosomes,Warframe,airsoft,bodybuilding,breakingmom,cars,cringe,eu4,exjw,explainlikeimfive,femalefashionadvice,food,heroesofthestorm,jailbreak,justneckbeardthings,oculus,pcmasterrace,photography,reddevils,relationships,rupaulsdragrace,stopdrinking,streetwear,techsupport,todayilearned,toronto,videos,xxfitness},
  yticklabels=none,
  xticklabel style={font=\sffamily\tiny, rotate=90},
  xtick=data,
  legend style={font=\sffamily\small},
    ]
    \nextgroupplot
    \confusionplot{#1}
    \nextgroupplot[colorbar right]
    \confusionplot{#2}
  \end{groupplot}
}

@@

\newcommand{\pplscatter}[2]{%
  \begin{tikzpicture}
  \begin{axis}[]
  \addplot[
    only marks, black,
  ] table [x=#1, y=#2] {floats/cclm_lmcc_ppl.csv};
  \end{axis}
  \end{tikzpicture}
}

>>>>

\newcommand{\commscatter}[3][\community]{%
  \addplot[
  scatter, only marks,
  font=\tiny,
  scatter src=explicit symbolic,
  scatter/classes={
    general-interest={mark=*, blue},
    videogames={mark=*, red},
    technology={mark=*, green},
    sports={mark=*, cyan},
    female-focused={mark=*, magenta},
    other={mark=o, black}%
  },
  nodes near coords*={#1},
  nodes near coords align={vertical},
  every node near coord/.append style={font=\sffamily\tiny},
  visualization depends on={value \thisrow{community} \as \community}
  ] table [x={#2}, y={#3}, meta=category] {floats/comm.csv};
}

\newcommand{\confusionplot}[1]{%
  \addplot[
  matrix plot*,
  shader=faceted,
  mesh/cols=46,
  point meta=explicit,
  ] table [x=actual_comm, y=pred_comm, meta=#1] {floats/confusion.csv};
}

\newcommand{\confusiongroupplot}[2]{%
  \begin{groupplot}[
    group style={group size=2 by 1}, 
    enlargelimits=false,
    axis line style={draw=none},
    %axis equal,
    xmin=-1, xmax=46,
    ymin=-1, ymax=46,
    width=8cm,
    xticklabels={Advice,AskWomen,BabyBumps,CFB,Drugs,EDH,EarthPorn,Fantasy,GameDeals,GlobalOffensive,Jokes,Kappa,KerbalSpaceProgram,KotakuInAction,LifeProTips,MLS,MMA,MaddenUltimateTeam,TwoXChromosomes,Warframe,airsoft,bodybuilding,breakingmom,cars,cringe,eu4,exjw,explainlikeimfive,femalefashionadvice,food,heroesofthestorm,jailbreak,justneckbeardthings,oculus,pcmasterrace,photography,reddevils,relationships,rupaulsdragrace,stopdrinking,streetwear,techsupport,todayilearned,toronto,videos,xxfitness},
    yticklabels=none,
    x tick label style={font=\sffamily\tiny,rotate=45,anchor=east},
    xtick=data,
    legend style={font=\sffamily\small},
    ]
    \nextgroupplot
    \confusionplot{#1}
    \nextgroupplot[colorbar right]
    \confusionplot{#2}
  \end{groupplot}
}

@@
